Porting of applications becomes very actual issue in the modern world of information technology. Every operating system
uses own interface and technologies for application development. However, there are cross-platform tools that allows
porting the applications without many modifications. One of~these tools is the Java programming language.

Java languge is developed by Oracle Corporation and it is distributed in several platforms. The most common platform is
Java Standart Edition which contains many of the~basic Java libraries ordinarily used in standard desktop applications.
A~set of~these libraries is called the Java SE API.

Android is an operating system for mobile devices developed by Google. It uses the~Java programming language to an
application development. Android runtime environment includes not only the libraries for development of user interface
of Android applications but~also subset of Java SE API libraries.

OptaPlanner is an open-source software developed by JBoss community designed for~solving planning problems. It is
completely writen in Java language and its easily portable between desktop operating systems. However, Android API does
not contain all of the Java SE API libraries and therefore porting of the OptaPlanner tool causes problems with
dependencies. This thesis deals with these problems and shows implemention of simple Android application which uses
OptaPlanner tool to solve the Vehicle routing problem.

Along with this introduction, this thesis is divided into another seven chapters. Chapter~\ref{JavaChapter} presents the
Java programing language and its platforms. Chapter~\ref{AndroidChapter} describes the Android platform, its
architecture and the build process. Chapter~\ref{OptaPlannerChapter} explains what the OptaPlanner tool is and shows how
can be used. In Chapter~\ref{PortingChapter}, differences between Java SE API and Android API are described and the
possible solutions of the JavaBeans problem on the Android platform are suggested. One of the solutions is selected and
used for realization of the portation. Design and implementation of sample Android Vehicle Routing Problem application
which uses OptaPlanner tool is presented in Chapter~\ref{ApplicationChapter}. Performed testings and future work is
described in Chapter~\ref{TestingChapter} and the last Chapter~\ref{ConclusionChapter} summarizes the entire work.

